### 2023-06-19 Draft.tex
Started separable filtering

\section{Introduction}
\section{Linear Filters}
With \textbf{Linear Filtering}, or specifically linear spatial filtering, the function which is used to pass the image through
must be linear and shift invariant.\smallskip

A common formula for linear filtering is the \emph{Correlation Filtering}. \cite{BovikActon09}
\[
    g(i,j) = \sum_{l \in \mathscr{M}}\sum_{k \in \mathscr{N}}{f(i+k, j+l) \cdot h(k, l)}
\]
or commonly notated as $g = f \otimes h$.\smallskip

@@

$$
\underbrace{
    \begin{bmatrix}
        128 & 34 & 123\\
        68 & 54 & 73 \\
        100 & 95 & 17
    \end{bmatrix}}_{\text{input neighborhood}}
\otimes
\underbrace{
    \begin{bmatrix}
        0.1 & 0.1 & 0.1\\
        0.1 & 0.2 & 0.1\\
        0.1 & 0.1 & 0.1
    \end{bmatrix}}_{\text{kernel}}
= 75
$$

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[correlation]{\includegraphics[width=0.19\textwidth]{img/correlation.jpg}}
    \subfigure[convolution]{\includegraphics[width=0.19\textwidth]{img/convolved}}
    \caption{Difference between Correlation and Convolution Filtering by translating 32 pixels}
    \label{fig:correlation_convolution}
\end{figure}\end{minipage}\smallskip

@@

There are other ways to handle the borders such as two-phase duplication,
which is similar to clamp but alternates between the last 2 pixels.
A change in the filter function per individual basis can also be implemented, but it might not be practical,
as large amounts of additional logic may be needed. \cite{BaileyAmbikumar18}

>>>>

\section{Introduction}
\section{Linear Filters}
With \textbf{Linear Filtering}, or specifically linear spatial filtering, the function which is used to pass the image through
must be linear and shift invariant. The function \pmb{L} is considered linear if there exists two constants $a$ and $b$ for any
two input pixels $f_1(m, n)$ and $f_2(m, n)$ such that

\begin{align*}
    \implies & g_1(m, n) = \pmb{L} [f_1(m, n)] \land g_2(m, n) = \pmb{L} [f_2(m, n)] \\
    \implies & a \cdot g_1(m, n) + b \cdot g_2(m, n) = \pmb{L} [a \cdot f_1(m, n) + b \cdot f_2(m, n)]
\end{align*}

also called the \emph{superposition property of linear systems}\cite{BovikActon09}. Whereas a function is considered \emph{shift invariant} when
\begin{align*}
    \implies & g(m, n) = \pmb{L} [f(m, n)]\\
    \implies & g(m - p, n - q) =  \pmb{L} [f(m - p, n - q)]
\end{align*} $p$ and $q$ implies a spatial shift to both the input and output pixels\cite{BovikActon09}. An intuitive way to think
of these two properties is that the function has to behave the same throughout the entire picture.\smallskip

\subsection{Correlation and Convolution Filtering}
A common formula for linear filtering is the \emph{Correlation Filtering}\cite{BovikActon09}. 
\[
    g(i,j) = \sum_{l \in \mathscr{M}}\sum_{k \in \mathscr{N}}{f(i+k, j+l) \cdot h(k, l)}
\]
or commonly notated as $g = f \otimes h$.\smallskip

@@

\[
\underbrace{
    \begin{bmatrix}
        128 & 34 & 123\\
        68 & 54 & 73 \\
        100 & 95 & 17
    \end{bmatrix}}_{\text{input neighborhood}}
\otimes
\underbrace{
    \begin{bmatrix}
        0.1 & 0.1 & 0.1\\
        0.1 & 0.2 & 0.1\\
        0.1 & 0.1 & 0.1
    \end{bmatrix}}_{\text{kernel}}
= 75
\]

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[correlation]{\includegraphics[width=0.19\textwidth]{img/correlation.jpg}}
    \subfigure[convolution]{\includegraphics[width=0.19\textwidth]{img/convolved}}
    \caption{Difference between Correlation and Convolution Filtering by translating 32 pixels}
\end{figure}\end{minipage}\smallskip

@@

There are other ways to handle the borders such as two-phase duplication,
which is similar to clamp but alternates between the last 2 pixels.
A change in the filter function per individual basis can also be implemented, but it might not be practical,
as large amounts of additional logic may be needed\cite{BaileyAmbikumar18}.

\subsection{Separable Filtering}
The time complexity to calculate a single pixel can be said to be $O(M \cdot N)$ as it is dependent
on the size of the kernel. For each of the pixel in the kernel, it must be both multiplied and summed together
with its neighbors. This can be in practice sped up by separating the horizontal and vertical operations, hence performing
a one dimensional convolution horizontally, then one additional operation for the vertical convolution. This kernel in this case is the
outer product of two one dimensional kernels\cite{Szeliski_2022}. 
\[
    \pmb{K} = v \otimes h = 
    \begin{bmatrix}
        v_1h_1 & v_1h_2 & v_1 h_3 & \hdots & v_1h_n \\
        v_2h_1 & v_2h_2 & v_2 h_3 & \hdots & v_2h_n \\
        \vdots & \vdots & \vdots & \ddots & \vdots \\
        v_m h_1 & v_m h_2 & v_m h_3 & \hdots & v_m h_n \\
    \end{bmatrix}
\]
